The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-31 Paper.Rmd
editted all

Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}). \textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data for species=16 [Name : "Ceratium furca"] from those locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}). \textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.

@@

The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulae.
AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation. 
The lower- and upper-tail dependence of the 
best-fitting member of each family was also provided by `BiCopEst`.

@@

\textbf{(i)} The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

@@

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data. Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. Therefore, we also verified this finding with \nameref{NPA} approach (see Fig. \ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam} contains a number indicating how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics. Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.

@@

We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
(i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations).

@@

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Total number of location pairs (excluding self-comparison)",
           "2. Number of location pairs having non-independent copula structures",
           "3. Number of non-Normal copula as best fit",
           "4. Average AIC-weights of best fit copula from all location pair",
           "5. Average AIC-weights of Normal copula from all location pair",
           "6. Number of good-fit location pair*",
           "7. 95% CI of model-avg. lower tail-dep. from all location pair",
           "8. 95% CI of model-avg. upper tail-dep. from all location pair",
           "9. Number of location pairs showing stronger lower tail dependence",
           "10. 95% CI of (Pl - Pu) from all location pair")
           )

@@

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (I) \hspace{3.2 cm} (J) \hspace{3.2 cm} (K) \hspace{3.2 cm} (L) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}
\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs. log(mass) for birds, \textbf{(C)} log(BMR) vs. log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. Lower panel shows \textbf{(I)} raw data , \textbf{(J)} copula plot for  green spruce aphid's count data from location pair(4,11) and \textbf{(K)} raw data , \textbf{(L)} copula plot for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}

@@

* First, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively. Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig. \ref{fig_Cause4copula_params_GOF_500}). We also tested the goodness of fit of the estimated parameters for the population-copula and its standard error due to sample size. We reported p-values from goodness of fit tests for both statistics (Cramer-von Mises and Kolmogorov-Smirnov). p-values greater or equal to $0.05$ (above the dashed line in Fig. \ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters. Fig. \ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig. \ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$. (For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise copula family. \label{fig_Cause4copula_params_GOF_500}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption{Effects of asymmetric global noise copula (the dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq. \ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

@@

This section is dedicated to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data. Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data. We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula. We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}). For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig. \ref{SM-fig_spTL}(I)).

@@

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged $(A)$ green-spruce aphid's abundance, $(B)$ leaf-curling plum aphid's firstflight date, $(C)$ ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation). \label{fig_skewness_spearman}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid's (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $10000$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data). $(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for leaf plum corn aphid's abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean). \label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}

@@

Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables. We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. These copulae were chosen carefully so that they covered the whole range from exclusively lower (Clayton, Survival Gumbel, Survival Joe, Survival BB6, Survival BB8) or upper tail dependence (Survival Clayton, Gumbel, Joe, BB6, BB8) to both tail dependence (BB1, Survival BB1, BB7, Survival BB7) in addition to two symmetric copulae (Normal and Frank). Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we innovated and we found they were consistent with the inferences we drew from model selction results. Our approach could help to interpret the synchronous behavior of a species from several metapopulations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongerly correlated than lower values.

>>>>

Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species. We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered. \textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}). \textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.

@@

The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`. Families considered (with abbreviations here indicated) included the **N**ormal/Gaussian, **C**layton, **G**umbel, **F**rank, **J**oe, **BB1**, **BB6**, **BB7**, **BB8**, **S**urvival **C**layton , **S**urvival **G**umbel, **S**urvival **J**oe, **S**urvival **BB1**, **S**urvival **BB6**, **S**urvival **BB7** and **S**urvival **BB8** copulae. These copulae were chosen carefully so that they covered the whole range from exclusively lower (C, SG, SJ, SBB6, SBB8) or upper tail dependence (SC, G, J, BB6, BB8) to both tail dependence (BB1, SBB1, BB7, SBB7) in addition to two symmetric copulae (N, F). AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation. 
The lower- and upper-tail dependence of the 
best-fitting member of each family was also provided by `BiCopEst`.

@@

\textbf{(i)} The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and sample variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

@@

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data. Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. Therefore, we also verified this finding with \nameref{NPA} approach (see Fig. \ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam} contains a number indicating how many Normal surrogates out of 1000, are different ($95\%$ confidence interval) from Normal-copula statistics. Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.

@@

We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other. Here, to present our results we chose one species from each category. We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets (i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations).

@@

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Total number of location pairs (excluding self-comparison)",
           "2. Number of location pairs having non-independent copula structures",
           "3. Number of non-Normal copula as best fit",
           "4. Average AIC-weights of best fit copula from all location pair",
           "5. Average AIC-weights of Normal copula from all location pair",
           "6. Number of good-fit location pair*",
           "7. 95% CI of model-avg. lower tail-dep. from all location pair",
           "8. 95% CI of model-avg. upper tail-dep. from all location pair",
           "9. Percentage of location pairs showing stronger lower tail dependence",
           "10. 95% CI of (Pl - Pu) from all location pair")
           )

@@

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
\end{center}
\textbf{ \hspace{7 cm} (I) \hspace{3.2 cm} (J)} \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=3.8 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
\includegraphics[width=3.8 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
\end{center}
\textbf{ \hspace{7 cm} (K) \hspace{3.2 cm} (L)} \\
\vspace{-0.35 cm}
\begin{center}
\includegraphics[width=3.8 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
\includegraphics[width=3.8 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs. log(mass) for birds, \textbf{(C)} log(BMR) vs. log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs. Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel. Lower panel shows \textbf{(I - J)} raw data plots and \textbf{(K - L)} copula plots for  green spruce aphid's count data from location pair(4,11) and for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}

@@

* First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively. Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig. \ref{fig_Cause4copula_params_GOF_500}). We also tested the goodness of fit of the estimated parameters for the population-copula and its standard error due to sample size. We reported p-values from goodness of fit tests for both statistics (Cramer-von Mises and Kolmogorov-Smirnov). p-values greater or equal to $0.05$ (above the dashed line in Fig. \ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters. Fig. \ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig. \ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$. (For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit (not significantly different) of population copula with same noise copula family. \label{fig_Cause4copula_params_GOF_500}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption{Effects of asymmetric global noise copula (resulting from the asymmetric dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq. \ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

@@

This section is dedicated to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We wanted to examine whether the asymmetric tail dependence had any **(i)** effects on skewness of spatially averaged time series data. We predicted that because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, should change significantly from the individual skewness of time-series data. We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}). For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig. \ref{SM-fig_spTL}(I)).

@@

\begin{figure}[!h]
\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
\vspace{-0.1 cm}
\begin{center}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid firstflight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation). \label{fig_skewness_spearman}}
\end{center}
\end{figure}

@@

\begin{figure}[!h]
\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $9961$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data). $(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for corn-leaf aphid abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean). \label{fig_sTL_mncurv_sp6}}
\end{center}
\end{figure}

@@

Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables. We applied two methods. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selction results. Our approach could help to interpret the synchronous behavior of a species from several metapopulations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations. Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean. Copulae could also give us information about the causal relationship between ecological variables. For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongerly correlated than lower values.
